State of the Practice for Lattice Boltzmann Method Software

### 2022-03-02 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
updates

\subsection{Augment Theory Manual to Include Requirements}

As discussed in Section~\ref{Sec_CompareArtifacts}, requirements specifications
are rarely written for research software. Although a full requirements
specification has advantages \citep{SmithEtAl2007, SmithAndLai2005},
given the lack of development time (\ref{P_LackDevTime}), we recommend the
intermediate step of augmenting the existing theory manuals with some
requirements information.  Table~\ref{artifactspresent} shows that theory
documentation is common.  We observed 17 of 24 packages had at least some theory
documentation. Templates for scientific requirements, like \citet{SmithEtAl2007,
SmithAndLai2005}, show theory is a significant part of the requirements
documentation. With the addition of some extra information, the theory documents
can be transformed into requirements specification.  The key extra information
includes explicitly stating user characteristics, explicitly stating how the
user interacts with the software in terms of input data requirements, and
listing likely and unlikely changes.  Explicit statements about likely changes
are invaluable in the design stage, since they provide developers guidance on
how general the software needs to be.  For those wishing to maximize the value
of a requirement specification, information could be added like prioritizing the
nonfunctional requirements (to show the relative importance between qualities
like portability, reliability and performance) and traceability information (to
show the consequences of changes to the assumptions). Spending time to
incorporate additional information in the theory manuals should help address the
quality of existing documentation (\ref{P_Documentation})  and improve the
software's usability (\ref{P_Usability}).

\subsection{Improve Repeatability, Reproducibility and Replicability}

- requirements are needed for the goal of replicability
\citep{BenureauAndRougier2017}.

- say what PPs are addressed
- record versions of everything
- Docker, Virtual machines

Repeatability - A correct output may not be exactly reproducible, as it may be
dependent on a probability distribution, so strictly comparing results may not
be appropriate. \wss{Ideas for dealing with reproducibility in these situations
will be explored in the recommendations section (like properties of the
solution, such as conservation of mass).  5R paper mentions repeatability -
design it so that the random number seed can be set.}

>>>>

\subsection{Augment Theory Manual to Include Requirements Information}

As discussed in Section~\ref{Sec_CompareArtifacts}, requirements specifications
are rarely written for research software. Although a full requirements
specification has advantages \citep{SmithEtAl2007, SmithAndLai2005},
given the lack of development time (\ppref{P_LackDevTime}), we recommend the
intermediate step of augmenting the existing theory manuals with some
requirements information.  Table~\ref{artifactspresent} shows that theory
documentation is common.  We observed 17 of 24 packages had at least some theory
documentation. Templates for scientific requirements, like \citet{SmithEtAl2007,
SmithAndLai2005}, show theory is a significant part of the requirements
documentation. With the addition of some extra information, the theory documents
can be transformed into requirements specification.  The key extra information
includes explicitly stating user characteristics, explicitly stating how the
user interacts with the software in terms of input data requirements, and
listing likely and unlikely changes.  Explicit statements about likely changes
are invaluable in the design stage, since they provide developers guidance on
how general the software needs to be.  For those wishing to maximize the value
of a requirement specification, information could be added like prioritizing the
nonfunctional requirements (to show the relative importance between qualities
like portability, reliability and performance) and traceability information (to
show the consequences of changes to the assumptions). Spending time to
incorporate additional information in the theory manuals should help address the
quality of existing documentation (\ref{P_Documentation})  and improve the
software's usability (\ref{P_Usability}).

\subsection{Improve Re-runnability, Repeatability, Reproducibility and Replicability}

- part of correctness (\ppref{P_Correctness})
- requirements are needed for the goal of replicability
\citep{BenureauAndRougier2017}.

	\item[Re-runnability] ``A re-runnable code is one that can be run again when
	needed, and in particular more than the one time that was needed to produce
	the results. It is important to notice that the re-runnability of a code is
	not an intrinsic property. Rather, it depends on the context, and becomes
	increasingly difficult as the code ages. Therefore, to be and remain
	re-runnable on other researchers' computers, a re-runnable code should
	describe---with enough details to be recreated---an execution environment in
	which it is executable.''  - record versions of everything - Docker, Virtual
	machines

	\item[Repeatability] Repeatability - A correct output may not be exactly
	reproducible, as it may be dependent on a probability distribution, so
	strictly comparing results may not be appropriate.  ``The code is running
	and producing the expected results. The next step is to make sure that you
	can produce the same output over successive runs of your program. In other
	words, the next step is to make your program deterministic, producing
	repeatable output. Repeatability is valuable. If a run of the program
	produces a particularly puzzling result, repeatability allows you to
	scrutinize any step of the execution of the program by re-running it again
	with extraneous prints, or inside a debugger. Repeatability is also useful
	to prove that the program did indeed produce the published results.
	Repeatability is not always possible or easy. But for sequential programs
	not depending on analog inputs, it often comes down to controlling the
	initialization of the pseudo- random number generators (RNG).''

	\item[Reproducibility] ``To recap, reproducibility implies re-runnability
	and repeatability and availability, yet imposes additional conditions.
	Dependencies and platforms must be described as precisely and as
	specifically as possible. Parameters values and inputs should accompany the
	result files. The data and scripts behind the graphs must be published. Unit
	tests are a good way to embed self-diagnostics of reproducibility in the
	code. Reproducibility is hard, yet tremendously necessary.''

	\item[Replicability] ``Replicability is the implicit assumption that any
	article that does not provide the code source makes: that the description it
	provides of the algorithms is sufficiently precise and complete to re-obtain
	the results it presents.'' Requires requirements.  For the current project,
	we did not assess the stronger form of ``reproducibility'' known as
	replicability.  A software project is considered replicable if the
	documentation alone (without the code) is sufficiently precise and complete
	to allow a third-part to re-obtain the results of the original code
	\citep{BenureauAndRougier2017}.  Replicability is difficult to achieve with
	current documentation practices.  This question should be considered in the
	next iteration of state of the practice assessments.

	As an example, \citet{IonescuAndJansson2013} summarize the problem that
	researchers had in reproducing the economic theory results obtained by Gintis in
	2006.  The scientific method requires that another researcher be able to
	independently replicate previous research results.
